---
title: "Multivariate analysis module case study: HIV spain"
author: "Niklas Willrich"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
    %\VignetteIndexEntry{Case_study_HIV}
    %\VignetteEngine{knitr::rmarkdown}
    %\VignetteEncoding{UTF-8}
---


This is the R version of original Stata-File

_this would be the place for the introduction to the case study_

## Loading packages and settings

```{r, warning=FALSE, message=FALSE, error=FALSE, comment=FALSE}
# Load packages
# require(foreign) # Read dta files version 12 and below
# require(readstata13) # Read dta files version 13 and higher
# require(dplyr) # Data manipulation
# require(survival) # Advanced regression analysis for survival analysis
# require(ggplot2) # Plotting

# Set the correct directory
# setwd("~/Documents/Projekte/epiet")

# Set knitr options to not evaluate anything
knitr::opts_chunk$set(eval=FALSE, warning=FALSE, message=FALSE, error=FALSE, comment=FALSE)
```


## Load data
```{r}
hiv.data <- read.dta("HIVexample2015.dta", convert.factors = FALSE)
```

## Generate the time variable ( NW: Simpler way to fill the variable conditionally ?)
```{r}
hiv.data$denom <- NA
hiv.data$denom[hiv.data$seroco == 0] <-
  (hiv.data$last_neg[hiv.data$seroco == 0]-
     hiv.data$first_test[hiv.data$seroco == 0])/365.25
hiv.data$denom[hiv.data$seroco == 1] <-
  (hiv.data$first_pos[hiv.data$seroco == 1]-
     hiv.data$first_test[hiv.data$seroco == 1])/365.25



# What about this way? I dont have the dataset, but I think it should work (Jakob)
hiv.data %>% 
    mutate(denom = ifelse(seroco==0, (last_neg - first_test)/365.25, (first_pos - first_test)/365.25 ))

```


## Prepare the data set for Poisson regression
```{r}
hiv.data.sum <- hiv.data %>%
  group_by(sex, idu, msm, sexwork, age2) %>%
  summarize(seroco.sum = sum(seroco), denom.sum = sum(denom))
```


## Save the data in a file
```{r}
save.image("HIVexample2015poisson.Rsv")

# Why do you save it in a Rsv file and not a Rdata file (Jakob)
```



## TODO: NOT SOLVED YET
ir seroco idu denom

## Poisson regression
```{r}
poisson.reg <-
  glm(seroco.sum ~ idu , family = poisson(link = "log"),
      data = hiv.data.sum, offset = log(denom.sum))
exp(poisson.reg$coefficients[[2]]) # same result as for Stata case
poisson.reg
confint(poisson.reg)

ggplot(hiv.data.sum, aes(seroco.sum)) +
  geom_histogram(binwidth = 5, origin = 0,
                 right = FALSE, fill = "red")
# slightly different from the stata-version, because we cannot specify
# the values to be discrete.

poisson.reg.2 <- update(poisson.reg,
                        formula = seroco.sum ~ sex + age2 +sexwork)

poisson.reg.3 <-
  glm(seroco.sum ~ sex + age2 + idu , family = poisson(link = "log"),
      data = hiv.data.sum, offset = log(denom.sum))

poisson.reg.4 <- update(poisson.reg,
                        formula = seroco.sum ~ sex + age2 + sexwork + idu + msm )

poisson.reg.5 <- update(poisson.reg,
                        formula = seroco.sum ~ sex + age2 + sexwork + idu )

poisson.reg.6 <- update(poisson.reg,
                       formula = seroco.sum ~ sex + age2 + sexwork + msm  )

poisson.reg.7 <- update(poisson.reg,
                        formula = seroco.sum ~ sex + age2 + idu + msm )

poisson.reg.7 <-
  glm(seroco.sum ~ sex + age2 + idu*sex + msm , family = poisson(link = "log"),
      data = hiv.data.sum, offset = log(denom.sum))

```


## TODO find nice summary functions to generate tables for several models
poisson.reg.7$fitted.values


## Plotting data
```{r}
cols <- c("Observed values"="#f04546","Predicted values"="#3591d1")
ggplot(hiv.data.sum) +
  geom_histogram(aes(seroco.sum, fill = "Observed values"),
                 binwidth = 5, origin = 0,
                 alpha = 0.5) +
  geom_histogram(aes(poisson.reg.7$fitted.values, fill = "Predicted values"),
                 binwidth = 5, origin = 0,
                 alpha = 0.5)
```


## Cox regression
```{r}
cox.reg <- coxph(Surv(denom, seroco) ~ idu+sex, data = hiv.data)
cox.reg.2 <- coxph(Surv(denom, seroco) ~ idu*msm + age + factor(centrocod),
                 data = hiv.data)

cox.reg.strata <- coxph(Surv(denom, seroco) ~ idu*msm + age + strata(centrocod),
                        data = hiv.data)

```


## Test of proportional hazard assumption
```{r}
cox.zph(cox.reg.2)
```

